---
title: "Assignment 04"
author: "Grayson Chen"
format: html
embed-resources: true
editor: visual
execute:
  warning: false
  message: false
---

## Data

I am interested in the pattern of labor force participation (LPR) in mainland China, so I found this data set from the International Labour Organization: <https://rplumber.ilo.org/data/indicator/?id=EAP_DWAP_SEX_AGE_RT_A&lang=en&type=label&format=.xlsx>

According to ILO's definition, the LPR is the share of persons who are in the labour force as a percent of the working-age population. The labour force is the sum of all persons of working age who are employed and those who are unemployed.

## Initial Steps

To set up the data (load data set & packages, clean data):

```{r}
library(tidyverse)
library(ggplot2)
library(readxl) #my data set is in .xlsx format; load this package to read

ilodata <- read_excel(
"data/ILO_labor_force_participation_rate.xlsx"
)

glimpse(ilodata)

#narrow the data down to "China" only; change the years into numeric values, and drop NA observations for the numeric columns.
chn <- ilodata |>
  filter(area == "China") |>
  mutate(year = as.numeric(time)) |>
  drop_na(year, obs_value)
glimpse(chn)
```

## Graph 1

I would like to find out the trend of total labor participation rate in 21th century. Data from "Official Estimate" contain year-level data since 2001.

```{r}
chnttl <- chn |>
  filter(source.label == "OE - Official Estimates", sex == "Total")
ggplot(chnttl,
       aes(x=year, y= obs_value)) +
  geom_line(color = "red", linewidth = 1.1) +
  geom_point(color = "red", size = 2) +
  labs(title = "Labor Force Participation Rate in China (2001–2018)",
       subtitle = "Total labor force participation among the population aged 15+",
       x = "Year",
       y = "Total Labor Participation Rate (%)",
       caption = "Source: ILOSTAT, Official Estimates")
  
```

#Observation This figure shows the trend in China's total labor force participation rate between 2001 and 2018. The participation rate declined steadily from more than 76 percent in the early 2000s to approximately 70 percent by 2018, indicating that a smaller share of the population aged 15 and above was participating in the labor market over time.

## Graph 2

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

## Graph 3

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

## Graph 4

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```
